---
title: "Analysis for Baccara paper on the interactions between the effects of climate and browsing on tree regeneration"
output: pdf_document
geometry: margin=1in
fontsize: 11pt
colorlinks: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Last compilation : \today


\vspace{20pt}

\begin{center}

Julien BARRERE

\end{center}

\vspace{20pt}


All the analyses presented here are available at the following [git](https://github.com/jbarrere3/BaccaraPaper) repository. 


# Effect of climate on browsing probability

The effect of climate on browsing probability ($p_{br}$) is fitted with the following model: 

$logit(p_{br}) = \alpha_1 + \alpha_2*T_{winter} + \alpha_3*Ht + \alpha_4*T_{winter}*Ht + \alpha_5*map + \alpha_4*map*Ht + \theta_{S/Z} + \epsilon$

Where $T_{winter}$ is the mean winter temperature, $map$ is the mean annual precipitation, $\alpha$ are parameters to estimate, $Ht$ is seedling height, and $\theta_{S/Z}$ is a random intercept with elevation (Z) nested in site (S). 

Before fitting the model, we verified that the two climatic variables were not too correlated: 

```{r fig_thivvsmap, echo=FALSE, out.width = '50%'}
knitr::include_graphics("fig/map_vs_thiv.pdf")
```


\newpage

We found that $T_{winter}$ had a negative effect on browsing probability for all species but Norway spruce, while precipitation affected browsing probability for both maple and spruce:  


```{r figBrowsing_hiv, echo=FALSE, fig.cap="Effect of height and (a) winter temperature or (b) precipitation on browsing probability, when all climatic variables are included in a single model. (c) estimate and confidence interval of the effect of each predictor.", out.width = '90%'}
knitr::include_graphics("fig/browsing_hiv_noG.pdf")
```

We also fitted a second model including ungulate mass density index (umdi) as an additional explanatory variable: 

$logit(p_{br}) = \alpha_1 + \alpha_2*T_{winter} + \alpha_3*Ht + \alpha_4*T_{winter}*Ht + \alpha_5*map + \alpha_4*map*Ht + \alpha_5*umdi + \theta_{S/Z} + \epsilon$

The umdi was calculated by summing in each site the density of each ungulate species (predicted by [ENETWILD et al. 2022](https://zenodo.org/record/7214870)) weighted by their body mass (retrieved from [naturalis database](https://naturalis.github.io/trait-organismal-ungulates/data/)). 


The parameters estimate, p-values, and VIF (to check for collinearities) of each model are shown in the table below: 

\newpage


```{r resultsBrowsing}
library(targets)
tar_load(browsing_models_hiv_umdi)
source("R/functions_data.R")
format_browsing_models(browsing_models_hiv_umdi, index.name = "umdi")
```


\newpage

# Effect of climate and browsing on seedling growth

We fitted the following model to study the effect of browsing ($Br$) and climate on seedling last shoot length (LSL)

$log(LSL) = \alpha_1 + \alpha_2*Ht + \alpha_3*Br \alpha_4*T_{mean} + \alpha_5*T_{mean}*Br + \alpha_6*map + \alpha_7*map*map + \theta_{S/Z} + \epsilon$

Where $T_{mean}$ is the mean annual temperature. 


```{r figGrowth, echo=FALSE, fig.cap="Effect of height, browsing and (a) temperature or (b) precipitation on seedling growth. (c) estimate and confidence interval of the effect of each predictor.", out.width = '90%'}
knitr::include_graphics("fig/growth_noG.pdf")
```


\newpage

The parameters estimate, p-values, and VIF (to check for collinearities) of each model are shown in the table below: 


```{r resultsGrowth}
library(targets)
tar_load(growth_models_noG)
source("R/functions_data.R")
format_growth_models(growth_models_noG)
```

